### 2018-04-10 paper/overview.tex
added diagram
pass on overview

%first introduce the question
In \autoref{sec:running-ex}, we describe a recursion-free CHC system,
$\mcchc$, that formulates verifying the safety of the program
\cc{multiCall} (\autoref{fig:multicall-code}).
%
In \autoref{sec:solve-ex}, we show that $\mcchc$ is a
\emph{Clause-Dependent Disjoint (CDD)} system and how \sys can solve it by 
encoding it into binary interpolants.
%
In \autoref{sec:not-in}, we illustrate that
$\mcchc$ is not in the directly solvable classes introduced in
previous work.
%

% introduce running example program
\subsection{Verifying \cc{multiCall}: an example hierarchical program}
\label{sec:running-ex}

\QZ{remake the graph}
% include code of running example and some figure:
\begin{figure}[t]
  \centering
  \begin{floatrow}[2]
    \ffigbox[.26\textwidth] %
    { \caption{\cc{multiCall}: an example hierarchical program.} %
      \label{fig:multicall-code} }
    { \input{code/multiCall.java} }
    %
    \ffigbox[.7\textwidth] %
    {\caption{Dependencies of CHC system $\mcchc$, depicted as a
        hypergraph.
        % 
        Each relational predicate is a node (circle) and each clause
        is a hyperedge (box). } %
      \label{fig:ex-deps} }
    { \includegraphics[width=\linewidth]{fig/ex-deps.pdf} }
  \end{floatrow}
\end{figure}
% walk through code of running example:
%
\cc{multiCall} includes a procedure \cc{db} that, given an integer
variable \cc{m}, returns the value of \cc{m} doubled (line 1).
%
\cc{main}, the entry procedure of \cc{multiCall}, given integer \cc{n}
(line 3), first binds variable \cc{abs} to $0$ (line 4).
%
\cc{main} tests if \cc{n} is greater than or equal to $0$ (line
5).
%
If so, \cc{main} binds the value of \cc{n} to \cc{res} (line 6).
%
If not, \cc{main} binds the negation value of \cc{n} to \cc{res} (line 8).
%
\cc{main} then calles \cc{dbl} on \cc{abs} and binds the return value to \cc{res} (line 9)
%
\cc{main} then returns the value stored in \cc{res} (line 10).

% state the safety problem:
The property to be verified in \cc{multiCall} is, for each input, 
\cc{multiCall}
returns a value greater than or equal to $0$.
% talk about solving Horn Clauses:
Verifying that \cc{multiCall} satisfies such a property can be reduced
to solving a recursion-free CHC system over a set of uninterpreted predicates
that each represents a control location and
procedure name in \cc{multiCall}.
%
In particular, one such system $\mcchc$ is
%
\QZ{left allign}
\begin{gather}
\label{chc}
  % semantic constraint of dbl:
  \cc{db}(\cc{m}, \cc{d}) \gets \cc{d} = 2 * \cc{m} \\
  % semantic constraint of line 3:
   \cc{L4}(\cc{n}, \cc{abs}) \gets  \cc{abs} = 0 \\
  % semantic constraint of line 5 (then branch)
   \cc{L6}(\cc{n}, \cc{abs}) \gets \cc{L4}(\cc{n}, \cc{abs}) \land \cc{n} \ge 0 \\
  % semantic constraint of line 6 (else branch)
   \cc{L8}(\cc{n}, \cc{abs}) \gets \cc{L4}(\cc{n}, \cc{abs}) \land \cc{n} < 0 \\
  % assign branch
   \cc{L9}(\cc{n}, \cc{abs'}) \gets \cc{L6}(\cc{n},\cc{abs}) \land \cc{abs'} = \cc{n} \\
   \cc{L9}(\cc{n}, \cc{abs'}) \gets \cc{L8}(\cc{n},\cc{abs}) \land \cc{abs'} = -\cc{n} \\
  % semantic constraint of main procedure:
    \cc{main}(\cc{n},\cc{res}) \gets \cc{L9}(\cc{n}, \cc{abs'}) \land \cc{db}(\cc{m}, \cc{d})
    \land \cc{abs'} = \cc{m} \land \cc{res} = \cc{d} \\
  %
   False \gets \cc{main}(\cc{n}, \cc{res}) \land \cc{res} < 0 
\end{gather}
%
While $\mcchc$ has been presented as the result
of a translation from hierarchical program \cc{multiCall}, \sys is
purely a solver for CHC systems: it does not require access to the
concrete representation of a program, or for a given CHC system to be
the result of translation from a program at all.

% solving the running example
\subsection{\cc{multiCall} as a Clause-Dependent Disjoint system}
\label{sec:solve-ex}
% state it is CDD
The recursion-free CHC system \cc{multiCall} is a \emph{Clause-Dependent-Disjoint (CDD)}) system.
% state the defn of CDD:
\emph{Clause-Dependent-Disjoint (CDD)}) system is an recursion-free CHC system such that
for any uninterpreted predictes in the body of the same clause, their transitive dependencies
are disjoint(i.e,they do not have common predicates).
%
The formal definition of CDD and its key properties are given in
\autoref{sec:CDD-defn}.

% give a example
In particular, for clause $(7)$, uninterpreted predicates \cc{L6} and \cc{db} shows in
the body of same clause.
%
The transitive dependency of \cc{L6} is \{\cc{L6}\,\cc{L8}\,\cc{L4}\}, 
and the transitive dependency of \cc{db} is $\varnothing$.
%
Their transitive dependencies are disjoint, because \{\cc{L6}\,\cc{L8}\,\cc{L4}\} $\cap \varnothing = \varnothing$.
%
The formal definition of transitive dependency is given in\QZ{reference}.
%
And the other clauses in \cc{multiCall} only has one uninterpreted predicates in the body, so
\cc{multiCall} is a CDD system.
%
The fact that $\mcchc$ is CDD corresponds to the fact that there no statements executed more than
once within one execution path.
%

% restate defn of CDD systems:
\sys solves $\mcchc$ directly by issuing a binary 
interpolation queriey for each uninterpreted predicates in $\mcchc$ in a topological order.
%
For each uninterpreted predicate $P$ in $\mathcal{S}$, the solution of
$P$ is the interpolant of two formulas: %
\textbf{(1)} a \emph{pre}-formula, constructed as 
the solution of each uninterpreted predicate in dependence of $P$, constraints
in corresponding clauses and boolean indicator variable for 
each uninterpreted predicates in $\mcchc$%
\textbf{(2)} a \emph{post}-formula, constructed as  
all constraints that occur after $P$ and boolean indicator variable for 
each uninterpreted predicates in $\mcchc$.
%

% walk through the example
E.g., when \sys solving $\mcchc$, \sys issues interpolation queries of 
each uninterpreted predicates in the order \cc{L4}, \cc{L6}, \cc{L8}, 
\cc{L9}, \cc{db} and \cc{main}.
%
In particular, when \sys issues interpolation query for uninterpreted predicates \cc{L9},
\sys has generated solutions for \cc{L4}, \cc{L6}, \cc{L8} which shows in\QZ{refer to graph}.
%

The solution of \cc{L9} is an interpolant of \textbf{(1)} 
a \emph{pre}-formula, constructed as solutions generated for \cc{L4}, \cc{L6}
the constraints in clauses (5) and (6) and boolean variables%
\textbf{(2)} a \emph{post}-formula constructed from constraints in
$\mcchc$ that all constrains after \cc{L9}, constrains for \cc{db}
and boolean variables.
%
In particular, the \emph{pre}-formula for \cc{L9} is 
\begin{gather}
((b_{L8}\land abs'= -n) \lor (b_{L6} \land abs=n))\land (\neg b_{L6} \lor n \ge 0) \land (\neg b_{L8} \lor n<0)
\end{gather}
%
, $b_{L8}$ and $b_{L6}$ are boolean variables corresponding to predicates \cc{L8} and \cc{L6}.
%
The \emph{post}-formula for \cc{L9} is
\begin{gather}
(b_{main} \land res' < 0 ) \land (\neg b_{main} \lor (b_{db} \land abs'= m \land res =d )) 
\land (\neg b_{db} \lor d = 2*m)
\end{gather}
, $b_{main}$ and $b_{db}$ are boolean variables corresponding to predicates \cc{main} and \cc{db}.
%
A detailed and formal description for the procedure of solving a CDD system is given in
\autoref{sec:solve-cdd}.

\subsection{\cc{multiCall} is not in other system}
\label{sec:not-in}
%
There are three classes of directly solvable systems has been introduced
in previous work(as described in \autoref{sec:intro}).
%
However, $\mcchc$ does not belong to any of these three classes.
%

@@

%linear systems:
Previous work also introduced linear system~\cite{albarghouthi12a}, 
which each clause's body has at most one uninterpreted predicate.
%
However, such systems cannot directly model the control flow of a
program that contains procedure calls.
%
$\mcchc$ is not a linear system because the body of clause (7) has two
predicates \cc{L9} and \cc{db}.
%
CHC solvers that solve linear systems, $\mcchc$, effectively inline constraints on
relational predicates that occur in non-linear clauses~\cite{albarghouthi12b}.
%
Such approaches can, in general, generate an inlined system that is
exponentially larger than a given system.

%% talk about expansion:
The class of CDD systems strictly contains these three classes, 
but does not contain all
recursion-free systems.
%
E.g., a recursion-free system that models \cc{multiCall} 
call \cc{dbl} twice on \cc{abs} in sequence would not
be CDD.
%
\sys, given an arbitrary recursion-free CHC system $S$,
solves it by generating an equivalent CDD system $S'$ and
solving $S'$ directly.
%
In general, $S'$ may have size exponential in the size of
$S$.
%
In practice, a suitable equivalent system can be generated using
heuristics analogous to those used to generate compact verification
conditions of hierarchical programs~\cite{flanagan01,lal-qadeer15}.
%
A general procedure for constructing a CDD expansion of a given CHC
system is given in \autoref{app:cons-cdd}.

>>>>

%first introduce the question
In \autoref{sec:running-ex}, we describe a recursion-free CHC system,
$\mcchc$, that formulates verifying the safety of a small program,
\cc{dblAbs} (\autoref{fig:multicall-code}).
%
In \autoref{sec:solve-ex}, we show that $\mcchc$ is a
\emph{Clause-Dependent Disjoint (CDD)} system and how \sys can solve it by
encoding it into binary interpolants.
%
In \autoref{sec:not-in}, we illustrate that $\mcchc$ is not in
directly solvable classes introduced in previous work.
%

% introduce running example program
\subsection{Verifying \cc{dblAbs}: an example hierarchical program}
\label{sec:running-ex}

% include code of running example and some figure:
\begin{figure}[t]
  \centering
  \begin{floatrow}[2]
    \ffigbox[.26\textwidth] %
    { \caption{\cc{dblAbs}: an example hierarchical program.} %
      \label{fig:multicall-code} }
    { \input{code/dblAbs.java} }
    %
    \ffigbox[.7\textwidth] %
    {\caption{$\mcchc$ shown as a directed hypergraph.
        % 
        Each relational predicate is depicted as a graph node
        while each clause is represented by a hyperedge.
        Each hyperedge is labelled by the fact in the corresponding CHC\@.
        Each node has a valid corresponding interpretation,
        written in braces.
      } %
      \label{fig:ex-graph} }
    % { \includegraphics[width=\linewidth]{fig/ex-deps.pdf} }
      {\input{fig/example}}
  \end{floatrow}
\end{figure}
% walk through code of running example:
%
\cc{dblAbs} is a small procedure that doubles the absolute value of
its input and stores the result in \cc{res}.
% state the safety problem:
The program also asserts that \cc{res} is greater than or equal to $0$
before exiting.
% talk about solving Horn Clauses:
Verifying this assertion can be reduced to solving a recursion-free
CHC system over a set of uninterpreted predicates that each represents
a control location and procedure name in \cc{dblAbs}.
%
In particular, one such system $\mcchc$ is
%
\begin{align}
\label{chc}
  % semantic constraint of dbl:
\cc{dbl}(\cc{x}, \cc{d}) &\gets \cc{d} = 2 * \cc{x} \\
  % semantic constraint of line 3:
  \cc{L}_4(\cc{n}, \cc{abs}) &\gets  \cc{abs} = 0 \\
  % semantic constraint of line 5 (then branch)
  \cc{L}_6(\cc{n}, \cc{abs}) &\gets \cc{L}_4(\cc{n}, \cc{abs}) \land \cc{n} \ge 0 \\
  % semantic constraint of line 6 (else branch)
  \cc{L}_8(\cc{n}, \cc{abs}) &\gets \cc{L}_4(\cc{n}, \cc{abs}) \land \cc{n} < 0 \\
  % assign branch
  \cc{L}_9(\cc{n}, \cc{abs'}) &\gets \cc{L}_6(\cc{n},\cc{abs}) \land \cc{abs'} = \cc{n} \\
   \cc{L}_9(\cc{n}, \cc{abs'}) &\gets \cc{L}_8(\cc{n},\cc{abs}) \land \cc{abs'} = -\cc{n} \\
  % semantic constraint of main procedure:
   \cc{main}(\cc{n},\cc{res}) &\gets \cc{L}_9(\cc{n}, \cc{abs'}) \land
                                     \cc{dbl}(\cc{x}, \cc{d})
    \land \cc{abs'} = \cc{x} \land \cc{res} = \cc{d} \\
  %
    \bot &\gets \cc{main}(\cc{n}, \cc{res}) \land \cc{res} < 0 
\end{align}
%
While $\mcchc$ has been presented as the result
of a translation from hierarchical program \cc{dblAbs}, \sys is
purely a solver for CHC systems: it does not require access to the
concrete representation of a program, or for a given CHC system to be
the result of translation from a program at all.

% solving the running example
\subsection{\cc{dblAbs} as a Clause-Dependent Disjoint system}
\label{sec:solve-ex}
% state it is CDD
The recursion-free CHC system $\mcchc$ is a
\emph{Clause-Dependent-Disjoint} (CDD) system.
% state the defn of CDD:
A CDD system is a recursion-free CHC system such that
each clause is dependence disjoint.
A clause is dependence disjoint if no two predicates in the body share
any transitive dependencies (i.e, they have no common predicates) and
there are no two predicates that are the same.
%
The formal definition of CDD and its key properties are given in
\autoref{sec:CDD-defn}.

% give a example
In particular, clause $(7)$ uses both $\cc{L}_9$ and $\cc{dbl}$ in its
body.
%
The transitive dependency of $\cc{L}_9$ is the set
$\{\cc{L}_4,\cc{L}_6,\cc{L}_8\}$, while the transitive dependency of
\cc{dbl} is the empty set.
%
Therefore, their transitive dependencies are disjoint:
$\{\cc{L}_4,\cc{L}_6,\cc{L}_8\} \cap \varnothing = \varnothing$.
If there were a conflict in the transitive dependencies, then $\mcchc$
would not be a CDD system.
%
The formal definition of transitive dependency is given in\QZ{reference}.
%
All other clauses in $\mcchc$ have at most one uninterpreted predicate
in the body, so they are trivially disjoint dependent.
%
Since every rule in $\mcchc$ is disjoint dependent, $\mcchc$ is a CDD
system.
%
The fact that $\mcchc$ is CDD corresponds to the fact that there no
statements executed more than once within one execution path.
\DAH{elaborate}
%

% restate defn of CDD systems:
\sys solves $\mcchc$ directly by issuing a binary interpolation query
for each uninterpreted predicate in topological order.
%
Each uninterpreted predicate $P$ has a valid interpretation which can
be computed by finding the interpolant of two formulas:
\textbf{(1)} the \emph{pre}-formula, constructed from the solutions of
each uninterpreted predicate $P$ directly depends on, constraints in
clauses where $P$ is the head, and boolean indicator variables for
each uninterpreted predicate in $\mcchc$%
\textbf{(2)} the \emph{post}-formula, constructed from
all constraints that occur after $P$ and boolean indicator
variables for each uninterpreted predicate in $\mcchc$.
%

We will walk through the interpolation query for $\cc{L}_9$ by hand.
By the time \sys reaches $\cc{L}_9$ it will already have solutions for
$\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$. Possible interpretations of these
predicates are shown in \autoref{fig:ex-graph}.
%
\DAH{WORK FROM HERE}
A solution for \cc{L9} can be constructed by interpolating \textbf{(1)}
the \emph{pre}-formula, constructed as solutions generated for \cc{L4}, \cc{L6}
the constraints in clauses (5) and (6) and boolean variables%
\textbf{(2)} the \emph{post}-formula constructed from constraints in
$\mcchc$ that all constraints after \cc{L9}, constraints for \cc{dbl}
and boolean variables.
%
In particular, the \emph{pre}-formula for \cc{L9} is 
\begin{gather}
  ((b_{L_8}\land \cc{abs'}= -\cc{n})
    \lor (b_{L_6} \land \cc{abs}=\cc{n}))
  \land (\neg b_{L_6} \lor \cc{n} \ge 0)
    \land (\neg b_{L_8} \lor \cc{n}<0)
\end{gather}
%
, $b_{L8}$ and $b_{L6}$ are boolean variables corresponding to predicates \cc{L8} and \cc{L6}.
%
The \emph{post}-formula for \cc{L9} is
\begin{gather}
  (b_{main} \land \cc{res'} < 0 )
    \land (\neg b_{main} \lor (b_{dbl} \land \cc{abs'}= \cc{x} \land \cc{res}=\cc{d} ))
  \land (\neg b_{dbl} \lor \cc{d} = 2*\cc{x})
\end{gather}
, $b_{main}$ and $b_{dbl}$ are boolean variables corresponding to
predicates \cc{main} and \cc{dbl}.
%
A detailed and formal description for the procedure of solving a CDD system is given in
\autoref{sec:solve-cdd}.

\subsection{\cc{dblAbs} is not in other system}
\label{sec:not-in}
%
There are three classes of directly solvable systems has been introduced
in previous work(as described in \autoref{sec:intro}).
%
However, $\mcchc$ does not belong to any of these three classes.
%

@@

%linear systems:
Previous work also introduced linear system~\cite{albarghouthi12a}, 
which each clause's body has at most one uninterpreted predicate.
%
However, such systems cannot directly model the control flow of a
program that contains procedure calls.
%
$\mcchc$ is not a linear system because the body of clause (7) has two
predicates \cc{L9} and \cc{dbl}.
%
CHC solvers that solve linear systems, $\mcchc$, effectively inline constraints on
relational predicates that occur in non-linear clauses~\cite{albarghouthi12b}.
%
Such approaches can, in general, generate an inlined system that is
exponentially larger than a given system.

%% talk about expansion:
The class of CDD systems strictly contains these three classes, 
but does not contain all
recursion-free systems.
%
E.g., a recursion-free system that models \cc{dblAbs} 
call \cc{dbl} twice on \cc{abs} in sequence would not
be CDD.
%
\sys, given an arbitrary recursion-free CHC system $S$,
solves it by generating an equivalent CDD system $S'$ and
solving $S'$ directly.
%
In general, $S'$ may have size exponential in the size of
$S$.
%
In practice, a suitable equivalent system can be generated using
heuristics analogous to those used to generate compact verification
conditions of hierarchical programs~\cite{flanagan01,lal-qadeer15}.
%
A general procedure for constructing a CDD expansion of a given CHC
system is given in \autoref{app:cons-cdd}.
